---
title: "Practical Machine Learning : Course Projects"
author: "W Wocher"
date: "April 24, 2016"
output: html_document
---

#Executive Summary
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
The goal of your project is to predict the manner in which the 6 participants did the exercise from accelerometers on the   
- belt   
- forearm   
- arm   
- dumbell   

The training data for this project are available here:  
- https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv  

The test data are available here:  
- https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv  

#Synopsis
Create 2 sets of data from the training data.   
 
1  Try several different models and compare how accurate each of them are.  
2  Once a model is decided on, then run predictions on the training data based on splitting is up into 2 data sets.   
3  Load the data from the test file  
4  Apply the model to this test file    

#Exploratory Data Analysis

##Install Libraries and Set Seed
```{r install_libraries}
library(lattice)
library(ggplot2)
library(mlbench)
library(caret)
library(randomForest)
library(foreach)
library(iterators)
library(parallel)
library(doParallel)
set.seed(64)
```

##Load Data
```{r load_data}
setwd("D:/Coursera/08 Practical Machine Learning/Project")
training.pml <- read.csv(file="pml-training.csv",head=TRUE,sep=",")
testing.pml <- read.csv(file="pml-testing.csv",head=TRUE,sep=",")
```

## Eliminate columns
Many of the columns do not have a lot of data.   Remove columns that have NAs in them.
```{r eliminate_na1}
training.complete <- training.pml[sapply(training.pml, function(x) !any(is.na(x)))] 
```

## Columns for analysis
The goal is to use data from accelerometers on the belt, forearm, arm, and dumbell.
```{r column_names}
column_names <- c("classe",
				  grep("belt", colnames(training.complete), value = TRUE)
				, grep("forearm", colnames(training.complete), value = TRUE)
				, grep("arm", colnames(training.complete), value = TRUE)
				, grep("dumbell", colnames(training.complete), value = TRUE)
				 )
column_names <- unique(column_names)
```

## Subset the data based on the columns
The goal is to use data from accelerometers on the belt, forearm, arm, and dumbell.
```{r column_subset}
training.subset <- subset(training.complete, select=column_names)
```
The data set currently has  `r ncol(training.subset)` columns.  This includes the outcome and the predictors.   
Modeling can not handle categorical predictors with more than 53 categories.  
The following steps is meant to reduce the number of columns.  

## Convert the columns to numeric.
This step introduces NAs.  Eliminate all the colunns with NAs.
```{r warning = FALSE}
for(i in 2:ncol(training.subset)) {
	training.subset[,i] = as.numeric(as.character(training.subset[,i]))
}
training.subset <- training.subset[sapply(training.subset, function(x) !any(is.na(x)))] 
```
The data set currently has  `r ncol(training.subset)` columns.   

## Plot
Plot the different data sets to see if anything has drastically changed
```{r data_plot}
par(mfrow=c(1,1))
plot(training.pml$classe, col="blue", main="All Data From File")
plot(training.complete$classe, col="yellow", main="Only Columns with Complete Data From File")
plot(training.subset$classe, col="green", main="Subset of  Columns with Complete Data From File Defined")
par(mfrow=c(1,1))
```

#Modeling

## Split the data
```{r data_split}
inTrain <- createDataPartition(y=training.subset$classe, p=0.7, list=FALSE)
training <- training.subset[inTrain,]
testing <- training.subset[-inTrain,]
```

## Setup parallel processing
Configure the trainControl object to be used in the modeling.
```{r parallel_processing_start}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster)
fitControl <- trainControl(method = "cv",
                           number = 10,
                           allowParallel = TRUE)
```


## Develop training model based on rf
```{r rf_model}
training.fit.rf <- train(classe ~ ., method="rf", data=training, trControl = fitControl)
```

## Develop training model based on knn
```{r knn_model}
training.fit.knn <- train(classe ~ ., method="knn", data=training, trControl = fitControl)
```

## Develop training model based on rpart
```{r rpart_model}
library(rpart)
training.fit.rpart <- train(classe ~ ., method="rpart", data=training, trControl = fitControl)
```
  

# Collect resamples and compare models
```{r model_comparisons}
results <- resamples(list(RF=training.fit.rf, KNN=training.fit.knn, RPART=training.fit.rpart))
summary(results)
par(mfrow=c(1,1))
dotplot(results)
```

After comparing the models, Random Forest has the best results.  

#Predicting

## Prediction of the Training File : Training Set
```{r rf_training_prediction_training}
training.prediction.rf <- predict(training.fit.rf$finalModel, training, type = "class")
confusionMatrix(training$classe,training.prediction.rf)
```

## Prediction of the Training File : Test Set
```{r rf_training_prediction_testing}
testing.prediction.rf <- predict(training.fit.rf$finalModel, testing, type = "class")
confusionMatrix(testing$classe,testing.prediction.rf)
```

## Prediction of the Testing File
Load the Testing file and apply the model
```{r rf_testing_prediction}
testing.pml.prediction.rf <- predict(training.fit.rf, testing.pml)

testing.pml.prediction.rf
```

## De-register parallel processing cluster
```{r parallel_processing_stop}
stopCluster(cluster)
```

